Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-01-07 Manuscript/manuscript.Rmd
review

We first assessed performances on random forest models built on fully unbalanced training sets across each resolution, for each predictor type (Figure 5) [??? Why not distance?; !!!included a  more desciptive plot with distance ??? Where? Don't see it on Figure 5, which exists in two versions (?); !!!figure5_v2_allnoneperfs, will delete other figures after meeting if deemed appropriate]. When looking at the the conventional metrics used for assessment (accuracy and AUC), models that were not corrected for class imbalance appeared to perform reasonably well.  However, when we considered the F1-score and MCC, we observed a significant degradation in performance across both resolution and predictor type. This was due in part to each models' inability to correctly classify the genomic bins that overlapped with flanked TAD boundaries (true positives, Supplementary Table 3). Even without correcting for the class imbalance, we found that, within predictor types, models with distance type predictors consistently outperformed the other two predictor types across all performance metrics.

@@

When comparing re-sampling techniques it was found that performing 50 iterations of random under-sampling yielded the best performances for both F1-Scores and MCC metrics, across all resolutions and predictor types (Figure 6). Models using 50-iterations of the SMOTE algorithm performed comparitively similar to those using RUS, across all resolutions and predictor types (supplementary Figure 1A). However, it was observed that models using 50-iterations of random over-sampling performed substantially worse than both RUS and SMOTE (Supplementary Figure 1B). More importantly, there was a notable decrease in both accuracy and AUC metrics when performing random under-sampling compared to models with no class balancing. This was evidenced by the more balanced interplay between the componenets of the confusion matrices created when validating the models (Supplementary Table 4). That is, less weight was carried by the true negative cases. This feature was also observed for models using SMOTE and ROS (Supplementary Table 5). As was the case for models with no class-balancing, there was a clear improvement when performing RUS for models with distance-type predictors, across all resolutions.[??? Again, where is distance?; !!!included]

>>>>

We first assessed performances on random forest models built on fully unbalanced training sets across each resolution, for each predictor type (Figure 5). When looking at the the conventional metrics used for assessment (accuracy and AUC), models that were not corrected for class imbalance appeared to perform reasonably well.  However, when we considered the F1-score and MCC, we observed a significant degradation in performance across both resolution and predictor type. This was due in part to each models' inability to correctly classify the genomic bins that overlapped with flanked TAD boundaries (true positives, Supplementary Table 3). Even without correcting for the class imbalance, we found that, within predictor types, models with distance type predictors consistently outperformed the other two predictor types across all performance metrics.

@@

When comparing re-sampling techniques it was found that performing 50 iterations of random under-sampling yielded the best performances for both F1-Scores and MCC metrics, across all resolutions and predictor types (Figure 6). Models using 50-iterations of the SMOTE algorithm performed comparitively similar to those using RUS, across all resolutions and predictor types (supplementary Figure 1A). However, it was observed that models using 50-iterations of random over-sampling performed substantially worse than both RUS and SMOTE (Supplementary Figure 1B). More importantly, there was a notable decrease in both accuracy and AUC metrics when performing random under-sampling compared to models with no class balancing. This was evidenced by the more balanced interplay between the componenets of the confusion matrices created when validating the models (Supplementary Table 4). That is, less weight was carried by the true negative cases. This feature was also observed for models using SMOTE and ROS (Supplementary Table 5). As was the case for models with no class-balancing, there was a clear improvement when performing RUS for models with distance-type predictors, across all resolutions.
